Work Fragmentation in Developer Interaction Data

### 2016-08-02 paper_2.tex
Highlighted the new changes

\usepackage{graphics}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{theorem}
\usepackage{verbatim} % comentarios
\usepackage{amsmath}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
\usepackage{mwe}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{multirow}
\usepackage{bigstrut}

\usepackage{array}
\newcommand\RR[1]{\textbf{Romain #1}}

@@

\begin{document}
%
\runningheads{L. Cruz, H. Sanchez, V. Gonzalez, R. Robbes}{Work Fragmentation in Developer Interaction Data}

@@

\begin{abstract}
%\boldmath
Information workers and software developers are exposed to work fragmentation, an interleaving of activities and interruptions during their normal work day. Small-scale observational studies have shown that this can be detrimental to their work. In this paper, we perform a large-scale study of this phenomenon for the particular case of software developers performing software evolution tasks. Our study is based on several thousands interaction traces collected by Mylyn and the Eclipse Usage Data Collector. We observe that work fragmentation is correlated to lower observed productivity at both the macro level (for entire sessions), and at the micro level (around markers of work fragmentation); further, longer activity switches seem to strengthen the effect. These observations give ground for subsequent studies investigating the phenomenon of work fragmentation.
\end{abstract}

@@

From these research questions, we got the following findings and conclusions:
So, in short these are our findings and conclusions:
\begin{itemize}
	\item We found an inverse relationship between the number of interruptions and our productivity metrics.
	\item A high proportion of prolonged interruptions is more detrimental to productivity than short interruptions.
	\item The productivity 30 minutes around an interruption is lower than the average.
	\item There are two well-defined patterns around the interruption, depending if the effect afterwards was positive or negative.
	\item There is a peak of productivity after the recovery time that lasts until approximately 10 minutes before the next interruption.
	\item There is a difference between the activities performed after an interruption, depending if it was positive or negative.
\end{itemize}
%\RR{For each research question, add a sentence explaining why we look into it}

@@

\textbf{Selection events. } Mylyn's selection feature is tied to high-level selection events in the Package Explorer. It does not include events such as switching tabs or navigation within and between documents via structured navigation (e.g., goto definition).

@@

Mylyn data has another characteristic which is unusual for time series: the time of occurrence of the events is not periodic (see column \textit{StartTime} in Table \ref{tbl:sample_event}). That is to say, events occur with non-equidistant separation gaps between them. However, in a time series, the values must be evenly spaced and chronologically sorted. For this reason, we compressed the size of the multivariate time series, pulling apart all the interruptions as a new time series variable. Then, each time series value represents the interruption duration in this minute (Figure \ref{fig:trans}). This transformation to a time series is similar to the Codealike's focus metric, who creates a time series with the quantification of the user's activity per minute, but without the decay and smoothing functions \cite{CLQ15}.

@@

\begin{figure}
	\centering	
	\begin{minipage}[c]{0.30\linewidth}
		\small
		\centering
			\begin{tabular}{| p{2cm} |p{2cm} |} 
				\hline 
				stat & value \\  
				\hline 
				\hline 
				total & 4,284 \\
				min. & 2 min \\
				1st Qu. & 18 min\\
				median & 35 min \\
				mean & 48.91 min \\
				3rd Qu. & 62 min \\
				max. & 535 min\\
				sd & 102.36 min\\
				\hline

@@

An important reason to consider the edit ratio is that having an interruption requires doing activities that might increase the execution of selection events, reducing the edit events. For example reading the code, navigating around files and starting a debugging sessions are activities heavily loaded with selections events and are common activities after an interruption, according to observational studies \cite{PR11,PD10}. The edit ratio can tell us when there is a reduction of edit events due to (mainly) selection events.

\paragraph{Correlation analysis of the metrics}
%\RR{perhaps expand a bit}
The metrics that characterize the productivity are related to the amount of activity recorded. Thus there is a degree of correlation between them; we study this correlation to determine whether the metrics measure different aspects or not.  %weand could be possible to find a correlation between them, specially with the number of edits and selections per minute, and the edit ratio.

We did a correlation analysis (using Pearson's correlation coefficient) involving these three metrics (Table \ref{tbl:corr_analysis}, classifying the data between traces with or without interruptions. We found that the Edits/min and Selections/minute metrics have a degree of correlation ($r=0.25$ and $r=0.35$), but that it is relatively low; this is the expected result as they explain two different kinds of activities. We do find that the correlation is higher in traces with interruptions. The same happens between the Selections/minute and Edit ratio metrics with the data of both kind of traces ($r=-0.38$ and $r=-0.32$), this time being a negative correlation which is also an expected value, as the Edit ratio should decrease when the Selections/minute increases.

@@

@@

\subsection{Classification of events}
Mylyn events belong to two main categories: edition and selection. The UDC data do not have this, so in order to match the classification of Mylyn we inferred the type of every event manually. So, we do not base our classification on previous work and we are aware that a change or misinterpretation could affect the final results. However, the attributes of the events offer enough information to make an accurate inference in most of the cases.

Note that we reuse the same term of selection in the following text, as it was used in Mylyn data. Strictly speaking, the word "navigation" would be better suited to the variety of events present in the UDC data, but we keep the old term for consistency.
%To select the event type, we made a thorough analysis of the different events to classify them where they better fit. In comparison with Mylyn we didn't have a classification by default but the attributes can give us a very good approximation.

@@

\begin{figure}
	\centering	
	\begin{minipage}[c]{0.30\linewidth}
		\small
		\centering
		\begin{tabular}{| p{2cm} |p{2cm} |} 
			\hline 
			stat & value \\  
			\hline 
			\hline 
			total & 15,825 \\
			min. & 2 min \\
			1st Qu. & 11 min\\
			median & 23 min \\
			mean & 35.87 min \\
			3rd Qu. & 47 min \\
			max. & 1,345 min\\
			sd & 43.95 min\\
			\hline

@@

Beyond visual inspection, we also quantify the statistical and the practical significance of these observations. First, in the results with Mylyn, all the differences observed are statistically significant with very low p-values (see Table~\ref{tbl:p_value}) according to the Mann-Whitney U-test. This is not surprising, given the shape of the boxplots and the size of the samples. It is important to mention that the values from the effect size Tables (\ref{tbl:p_value} and \ref{tbl:p_value_udc}) are obtained by comparing the \textit{none} group with the rest of the groups.

@@

   & none & $\leq 3$ & $[4 - 5]$ & $[6 - 10]$ & $\geq 11$  \\  
  \hline
  \multicolumn{6}{c}{\textbf{Edits}} \\
  \hline
  median & 6.56 & 1 & 0.67 & 0.65 & 0.6	 \\
  \cline{2-6} 
  mean & 6.29 &	2.59 & 1.55 & 1.29 & 0.91  \\ 
  % \cline{3-6} 
  %t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
   \cline{3-6} 
  U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
  \cline{3-6} 
  Cliff's delta & $\hookrightarrow$	& \textbf{0.61} & \textbf{0.77} & \textbf{0.80} & \textbf{0.87}    \\
  \hline

  \multicolumn{6}{c}{\textbf{Selections}} \\
  \hline 
  median & 4.56 & 2.08 & 1.71 & 1.42 & 1.28 \\
  \cline{2-6} 
  mean & 4.73 &	2.93 & 2.26 & 1.79 & 1.54  \\ 
    % \cline{3-6} 
    %t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 8.1e-13 & \multicolumn{3}{c}{$<$ 4.4e-16} \\

@@

  \multicolumn{6}{c}{\textbf{Edit ratio}} \\
  \hline 
  median & 0.56 & 0.36 & 0.30 & 0.32 & 0.33 \\
  \cline{2-6} 
  mean & 0.55 & 0.39 & 0.33 & 0.34 & 0.33 \\ 
  \cline{3-6} 
%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 4.2e-14& \multicolumn{3}{c}{$<$ 2.9e-16} \\
    \cline{3-6} 
    Cliff's delta & $\hookrightarrow$ & \textbf{0.47} & \textbf{0.66} & \textbf{0.67} & \textbf{0.73} \\ 
\hline

@@

		& none & $\leq 2$ & $[3 - 4]$ & $[5 - 7]$ & $\geq 7$  \\  
		\hline
		\multicolumn{6}{c}{\textbf{Edits}} \\
		\hline
		median & 3.58 & 1.5 & 1.23 & 1.15 & 1.18 \\
		\cline{2-6} 
		mean & 3.97 &	2.18 & 1.82 & 1.70 & 1.59 \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.42} & \textbf{0.53} & \textbf{0.57} & \textbf{0.63}    \\

		\hline
		\multicolumn{6}{c}{\textbf{Selections}} \\
		\hline 
		median & 2.54 & 1.66 & 1.5 & 1.5 & 1.4 \\
		\cline{2-6} 
		mean & 3.10 &	2.93 & 1.80 & 1.77 & 1.63  \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
		\cline{3-6} 
		U-test (adjusted)& $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.29} & \textbf{0.38} & \textbf{0.39} & \textbf{0.49} \\  
		\hline

		\multicolumn{6}{c}{\textbf{Edit ratio}} \\
		\hline 
		median & 0.57 & 0.49 & 0.45 & 0.43 & 0.44 \\
		\cline{2-6} 
		mean & 0.68 & 0.56 & 0.55 & 0.54 & 0.57 \\ 
		\cline{3-6} 
		%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) &  $\hookrightarrow$& \multicolumn{4}{c}{$<$ 0.0001} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$ & \textbf{0.19} & \textbf{0.24} & \textbf{0.28} & \textbf{0.26} \\ 
		\hline

@@

We tested the statistical significance of this results and show the results in the Table \ref{tbl:effect_size_dur_mylyn} for Mylyn, and Table \ref{tbl:effect_size_dur_udc} for UDC. For every group according to the number of interruptions, we measured the effect between the low and high groups of sessions, with the Cliff's delta test and U-test. The groups with one or two interruptions have moderate size effects, in comparison with the other groups with more interruptions that have lower size effects; this matches our observations via visual inspection on a more defined difference between the low and high groups when having a small number of interruptions. Starting from (approximately) three interruptions, the distributions of the metrics are very alike, except in some cases like the edit ratio with seven interruptions in Mylyn, where the median is almost the same but the distribution of the values are different enough to have a medium effect size.

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Effect size and statistical significance of the relationships between the groups of sessions by proportion of prolonged interruptions and the number of interruptions in Mylyn} %, and editions and selections.}
	\label{tbl:effect_size_dur_mylyn}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & 5.6e-10 & 6.5e-07 & 2.9e-04 & 0.0074 & 0.0021 & 0.32 &  0.39\\
		\hline
		Cliff's delta & \textbf{0.36} & \textbf{0.26} & \textbf{0.23} & \textbf{0.16} & \textbf{0.26} & \textbf{0.08} & \textbf{-0.09} \\
		\hline

@@

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Effect size and statistical significance of the relationships between the groups of sessions by proportion of prolonged interruptions and the number of interruptions in UDC} %, and editions and selections.}
	\label{tbl:effect_size_dur_udc}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & 1.9e-20 & 4.0e-13 & 0.0011 & 0.027 & 7.2e-08 & 0.055 & 6.5e-06\\
		\hline
		Cliff's delta & \textbf{0.36} & \textbf{0.27} & \textbf{0.08} & \textbf{0.06} & \textbf{0.17} & \textbf{0.07} & \textbf{0.18} \\
		\hline

@@

We did so by computing Cohen's $d$ on the quantity of edits before and after the interruption. To obtain a significant effect, we need the presence of activity both before and after the interruption. Not all the interruptions meet this criteria however: some are located close to the start or the end of a session, or too close to another interruption.  In total, 53\% of the interruptions had 30 minutes before and after the interruption and were selected for the analysis.  The remaining 47\% of the interruptions can not be used on this analysis due to the lack of time around them. %We used the Algorithm \ref{alg:local_effect} for the classification, and the 
Table \ref{tbl:local_effect} shows the applied thresholds and the results, accompanied with a typical example of an interruption in each category. This local analysis shows that there are indeed three well-defined groups of interruptions, with the two largest of them having clear effects on the activity in the session. The 45\% of the interruptions are positive in Mylyn, and the 44\% in UDC. The negative interruptions constitute the 44\% in Mylyn and UDC as well. And the neutral interruptions are the 11\% for Mylyn and 12\% for UDC.

@@

We can see in the Figure \ref{fig:box_recovery_udc} the evolution of the metrics through
the phases. The edits per minute and edit ratio show a tendency to increase and reach a maximum between the last phase of the recovery time and the peak of the session. Moreover, the fall of the productivity on the last minutes before the next interruptions agrees with the observations on the local analysis of the time around an interruption. From this we can conclude that the first five minutes after an interruption and the moment before the next one are the are the phases where productivity is more affected, while the phases around the peak are where the productivity is the highest.

@@

In RQ3, we saw different patterns of activity around positive and negative interruptions. In particular, we observed that the edits per minute were higher after a positive interruption; on the contrary they were lower after a negative interruption. The effect is even more marked in the first 5 minutes after the interruption, as can be seen in the Figure \ref{fig:box_recovery_effect_udc}.

@@

A first challenge that we encountered was that, similarly to the earlier research questions, the high number of different commands made it hard to infer broader trends. We hence created an additional classification of the commands we encountered in the activity logs. In a similar manner to the first classification, we followed an iterative process involving two of the authors to classify the events in these categories. Differently to the first classification, we did not limit ourselves to a restricted number of categories: rather, we let the classification emerge from the iterative process. We ended up with 11 categories, as described in Table \ref{tbl:new_events_udc}. We assigned the label and description that better suited the events grouped on that category and the associated type that reflects the majority of those events.

@@

\begin{table}[ht!]
	\small
	\renewcommand{\arraystretch}{1.3}
	\caption{New category of events. }
	\label{tbl:new_events_udc}
	\centering
	\begin{tabular}{|p{2.5cm}|p{9 cm}| p{2cm} |}
		\hline 
		category & description & type \\  
		\hline 
		\hline 
		high-nav & High-level navigation through classes, windows and artifacts & navigation \\
		file & Management of files (saving, closing, opening, etc.) & navigation\\
		edit-text & Text edition events & edition \\
		debug & Events executed during debugging sessions & navigation \\
		text-nav & Low-level text navigation & edition \\
		search & Text or code search & navigation \\
		find-replace & Search and replacement of text or code entities & navigation\\
		perspective & Change of perspectives in Eclipse & navigation  \\
		control & Source control events & navigation \\
		refactoring & Refactoring events & edition \\
		clean-build &  Events related to the build system (clean, build or run a project)  & navigation \\
		\hline
	\end{tabular}
\end{table}

@@

The second group of interruptions seems to be more disruptive as they require more time to reach the average productivity, and the activities are a reflection of the tasks performed during the recovery time as mentioned by Parnin \cite{PR11}, and they can be matched to the categories high-nav, debug and search, which are the common categories after a negative interruption.

@@

>>>>

\usepackage{graphics}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{theorem}
\usepackage{verbatim} % comentarios
\usepackage{amsmath}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
\usepackage{mwe}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{multirow}
\usepackage{bigstrut}
\usepackage{array}

\newboolean{showchangedforreviewer}

\setboolean{showchangedforreviewer}{true} % toggle to show or hide sentences and paragraphs requested by the reviewers.

\ifthenelse{\boolean{showchangedforreviewer}} {

	\newcommand{\changedforreviewer}[1]{\textcolor{blue}{#1}}

	\newenvironment{changedforreviewerlong}%
	{\begin{color}{blue}}%

		{\end{color}}

}{

\newcommand{\changedforreviewer}[1]{#1}

\newenvironment{changedforreviewerlong}%

{}

{}

}

\newcommand\RR[1]{\textbf{Romain #1}}

@@

\begin{document}

%
\runningheads{L. Cruz, H. Sanchez, V. Gonzalez, R. Robbes}{Work Fragmentation in Developer Interaction Data}

@@

\begin{abstract}

%\boldmath
Information workers and software developers are exposed to work fragmentation, an interleaving of activities and interruptions during their normal work day. Small-scale observational studies have shown that this can be detrimental to their work. In this paper, we perform a large-scale study of this phenomenon for the particular case of software developers performing software evolution tasks. Our study is based on several thousands interaction traces collected by Mylyn and the Eclipse Usage Data Collector. We observe that work fragmentation is correlated to lower observed productivity at both the macro level (for entire sessions), and at the micro level (around markers of work fragmentation); further, longer activity switches seem to strengthen the effect. These observations give ground for subsequent studies investigating the phenomenon of work fragmentation.
\end{abstract}

@@

\begin{changedforreviewerlong}
From these research questions, we got the following findings and conclusions:
\begin{itemize}
	\item We found an inverse relationship between the number of interruptions and our productivity metrics.
	\item A high proportion of prolonged interruptions is more detrimental to productivity than short interruptions.
	\item The productivity 30 minutes around an interruption is lower than the average.
	\item There are two well-defined patterns around the interruption, depending if the effect afterwards was positive or negative.
	\item There is a peak of productivity after the recovery time that lasts until approximately 10 minutes before the next interruption.
	\item There is a difference between the activities performed after an interruption, depending if it was positive or negative.
\end{itemize}
%\RR{For each research question, add a sentence explaining why we look into it}
\end{changedforreviewerlong}

@@

\begin{changedforreviewerlong}
\textbf{Selection events. } Mylyn's selection feature is tied to high-level selection events in the Package Explorer. It does not include events such as switching tabs or navigation within and between documents via structured navigation (e.g., goto definition).
\end{changedforreviewerlong}

@@

\begin{changedforreviewerlong}
Mylyn data has another characteristic which is unusual for time series: the time of occurrence of the events is not periodic (see column \textit{StartTime} in Table \ref{tbl:sample_event}). That is to say, events occur with non-equidistant separation gaps between them. However, in a time series, the values must be evenly spaced and chronologically sorted. For this reason, we compressed the size of the multivariate time series, pulling apart all the interruptions as a new time series variable. Then, each time series value represents the interruption duration in this minute (Figure \ref{fig:trans}). This transformation to a time series is similar to the Codealike's focus metric, who creates a time series with the quantification of the user's activity per minute, but without the decay and smoothing functions \cite{CLQ15}.
\end{changedforreviewerlong}

@@

\begin{figure}
	\centering	
	\begin{minipage}[c]{0.30\linewidth}
		\small
		\centering
			\begin{tabular}{| p{2cm} |p{2cm} |} 
				\hline 
				stat & value \\  
				\hline 
				\hline 
				total & 4,284 \\
				min. & 2 min \\
				1st Qu. & 18 min\\
				median & 35 min \\
				mean & 48.91 min \\
				3rd Qu. & 62 min \\
				max. & 535 min\\
				\changedforreviewer{sd} & 102.36 min\\
				\hline

@@

\begin{changedforreviewerlong} 
An important reason to consider the edit ratio is that having an interruption requires doing activities that might increase the execution of selection events, reducing the edit events. For example reading the code, navigating around files and starting a debugging sessions are activities heavily loaded with selections events and are common activities after an interruption, according to observational studies \cite{PR11,PD10}. The edit ratio can tell us when there is a reduction of edit events due to (mainly) selection events.
\end{changedforreviewerlong}

\paragraph{Correlation analysis of the metrics}
%\RR{perhaps expand a bit}
\begin{changedforreviewerlong}
The metrics that characterize the productivity are related to the amount of activity recorded. Thus there is a degree of correlation between them; we study this correlation to determine whether the metrics measure different aspects or not.  %weand could be possible to find a correlation between them, specially with the number of edits and selections per minute, and the edit ratio. 
We did a correlation analysis (using Pearson's correlation coefficient) involving these three metrics (Table \ref{tbl:corr_analysis}, classifying the data between traces with or without interruptions. We found that the Edits/min and Selections/minute metrics have a degree of correlation ($r=0.25$ and $r=0.35$), but that it is relatively low; this is the expected result as they explain two different kinds of activities. We do find that the correlation is higher in traces with interruptions. The same happens between the Selections/minute and Edit ratio metrics with the data of both kind of traces ($r=-0.38$ and $r=-0.32$), this time being a negative correlation which is also an expected value, as the Edit ratio should decrease when the Selections/minute increases.

@@

\end{changedforreviewerlong}

@@

\subsection{Classification of events}
\begin{changedforreviewerlong}
Mylyn events belong to two main categories: edition and selection. The UDC data do not have this, so in order to match the classification of Mylyn we inferred the type of every event manually. So, we do not base our classification on previous work and we are aware that a change or misinterpretation could affect the final results. However, the attributes of the events offer enough information to make an accurate inference in most of the cases.

Note that we reuse the same term of selection in the following text, as it was used in Mylyn data. Strictly speaking, the word "navigation" would be better suited to the variety of events present in the UDC data, but we keep the old term for consistency.
\end{changedforreviewerlong}
%To select the event type, we made a thorough analysis of the different events to classify them where they better fit. In comparison with Mylyn we didn't have a classification by default but the attributes can give us a very good approximation.

@@

\begin{figure}
	\centering	
	\begin{minipage}[c]{0.30\linewidth}
		\small
		\centering
		\begin{tabular}{| p{2cm} |p{2cm} |} 
			\hline 
			stat & value \\  
			\hline 
			\hline 
			total & 15,825 \\
			min. & 2 min \\
			1st Qu. & 11 min\\
			median & 23 min \\
			mean & 35.87 min \\
			3rd Qu. & 47 min \\
			max. & 1,345 min\\
			\changedforreviewer{sd} & 43.95 min\\
			\hline

@@

\begin{changedforreviewerlong}
Beyond visual inspection, we also quantify the statistical and the practical significance of these observations. First, in the results with Mylyn, all the differences observed are statistically significant with very low p-values (see Table~\ref{tbl:p_value}) according to the Mann-Whitney U-test. This is not surprising, given the shape of the boxplots and the size of the samples. It is important to mention that the values from the effect size Tables (\ref{tbl:p_value} and \ref{tbl:p_value_udc}) are obtained by comparing the \textit{none} group with the rest of the groups.
\end{changedforreviewerlong}

@@

   & none & $\leq 3$ & $[4 - 5]$ & $[6 - 10]$ & $\geq 11$  \\  
  \hline
  \multicolumn{6}{c}{\textbf{Edits}} \\
  \hline
  \changedforreviewer{median} & 6.56 & 1 & 0.67 & 0.65 & 0.6	 \\
  \cline{2-6} 
  mean & 6.29 &	2.59 & 1.55 & 1.29 & 0.91  \\ 
  % \cline{3-6} 
  %t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
   \cline{3-6} 
  U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
  \cline{3-6} 
  Cliff's delta & $\hookrightarrow$	& \textbf{0.61} & \textbf{0.77} & \textbf{0.80} & \textbf{0.87}    \\
  \hline

  \multicolumn{6}{c}{\textbf{Selections}} \\
  \hline 
  \changedforreviewer{median} & 4.56 & 2.08 & 1.71 & 1.42 & 1.28 \\
  \cline{2-6} 
  mean & 4.73 &	2.93 & 2.26 & 1.79 & 1.54  \\ 
    % \cline{3-6} 
    %t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 8.1e-13 & \multicolumn{3}{c}{$<$ 4.4e-16} \\

@@

  \multicolumn{6}{c}{\textbf{Edit ratio}} \\
  \hline 
  \changedforreviewer{median} & 0.56 & 0.36 & 0.30 & 0.32 & 0.33 \\
  \cline{2-6} 
  mean & 0.55 & 0.39 & 0.33 & 0.34 & 0.33 \\ 
  \cline{3-6} 
%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
     \cline{3-6} 
    U-test (adjusted) & $\hookrightarrow$ & 4.2e-14& \multicolumn{3}{c}{$<$ 2.9e-16} \\
    \cline{3-6} 
    Cliff's delta & $\hookrightarrow$ & \textbf{0.47} & \textbf{0.66} & \textbf{0.67} & \textbf{0.73} \\ 
\hline

@@

		& none & $\leq 2$ & $[3 - 4]$ & $[5 - 7]$ & $\geq 7$  \\  
		\hline
		\multicolumn{6}{c}{\textbf{Edits}} \\
		\hline
		\changedforreviewer{median} & 3.58 & 1.5 & 1.23 & 1.15 & 1.18 \\
		\cline{2-6} 
		mean & 3.97 &	2.18 & 1.82 & 1.70 & 1.59 \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) & $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.42} & \textbf{0.53} & \textbf{0.57} & \textbf{0.63}    \\

		\hline
		\multicolumn{6}{c}{\textbf{Selections}} \\
		\hline 
		\changedforreviewer{median} & 2.54 & 1.66 & 1.5 & 1.5 & 1.4 \\
		\cline{2-6} 
		mean & 3.10 &	2.93 & 1.80 & 1.77 & 1.63  \\ 
		% \cline{3-6} 
		%t.test & $\hookrightarrow$& 6.258e-08 & 1.623e-12 & 4.87e-16 & $<$ 2.2e-16 \\
		\cline{3-6} 
		U-test (adjusted)& $\hookrightarrow$ & \multicolumn{4}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$	& \textbf{0.29} & \textbf{0.38} & \textbf{0.39} & \textbf{0.49} \\  
		\hline

		\multicolumn{6}{c}{\textbf{Edit ratio}} \\
		\hline 
		\changedforreviewer{median} & 0.57 & 0.49 & 0.45 & 0.43 & 0.44 \\
		\cline{2-6} 
		mean & 0.68 & 0.56 & 0.55 & 0.54 & 0.57 \\ 
		\cline{3-6} 
		%    t.test & $\hookrightarrow$& 1.62E-13 & \multicolumn{3}{c}{$<$ 2.2e-16} \\
		\cline{3-6} 
		U-test (adjusted) &  $\hookrightarrow$& \multicolumn{4}{c}{$<$ 0.0001} \\
		\cline{3-6} 
		Cliff's delta & $\hookrightarrow$ & \textbf{0.19} & \textbf{0.24} & \textbf{0.28} & \textbf{0.26} \\ 
		\hline

@@

We tested the statistical significance of this results and show the results in the Table \ref{tbl:effect_size_dur_mylyn} for Mylyn, and Table \ref{tbl:effect_size_dur_udc} for UDC. For every group according to the number of interruptions, we measured the effect between the low and high groups of sessions, with the Cliff's delta test and U-test. \changedforreviewer{The groups with one or two interruptions have moderate size effects}, in comparison with the other groups with more interruptions that have lower size effects; this matches our observations via visual inspection on a more defined difference between the low and high groups when having a small number of interruptions. Starting from (approximately) three interruptions, the distributions of the metrics are very alike, except in some cases like the edit ratio with seven interruptions in Mylyn, where the median is almost the same but the distribution of the values are different enough to have a medium effect size.

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{\changedforreviewer{Effect size and statistical significance of the relationships between the groups of sessions by proportion of prolonged interruptions and the number of interruptions in Mylyn}}
	\label{tbl:effect_size_dur_mylyn}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & 5.6e-10 & 6.5e-07 & 2.9e-04 & 0.0074 & 0.0021 & 0.32 &  0.39\\
		\hline
		Cliff's delta & \textbf{0.36} & \textbf{0.26} & \textbf{0.23} & \textbf{0.16} & \textbf{0.26} & \textbf{0.08} & \textbf{-0.09} \\
		\hline

@@

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{\changedforreviewer{Effect size and statistical significance of the relationships between the groups of sessions by proportion of prolonged interruptions and the number of interruptions in UDC}}
	\label{tbl:effect_size_dur_udc}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & 1.9e-20 & 4.0e-13 & 0.0011 & 0.027 & 7.2e-08 & 0.055 & 6.5e-06\\
		\hline
		Cliff's delta & \textbf{0.36} & \textbf{0.27} & \textbf{0.08} & \textbf{0.06} & \textbf{0.17} & \textbf{0.07} & \textbf{0.18} \\
		\hline

@@

We did so by computing Cohen's $d$ on the quantity of edits before and after the interruption. To obtain a significant effect, we need the presence of activity both before and after the interruption. Not all the interruptions meet this criteria however: some are located close to the start or the end of a session, or too close to another interruption.  In total, 53\% of the interruptions had 30 minutes before and after the interruption and were selected for the analysis.  The remaining 47\% of the interruptions can not be used on this analysis due to the lack of time around them. %We used the Algorithm \ref{alg:local_effect} for the classification, and the

\begin{changedforreviewerlong}
Table \ref{tbl:local_effect} shows the applied thresholds and the results, accompanied with a typical example of an interruption in each category. This local analysis shows that there are indeed three well-defined groups of interruptions, with the two largest of them having clear effects on the activity in the session. The 45\% of the interruptions are positive in Mylyn, and the 44\% in UDC. The negative interruptions constitute the 44\% in Mylyn and UDC as well. And the neutral interruptions are the 11\% for Mylyn and 12\% for UDC.
\end{changedforreviewerlong}

@@

\begin{changedforreviewerlong}
We can see in the Figure \ref{fig:box_recovery_udc} the evolution of the metrics through
the phases. The edits per minute and edit ratio show a tendency to increase and reach a maximum between the last phase of the recovery time and the peak of the session. Moreover, the fall of the productivity on the last minutes before the next interruptions agrees with the observations on the local analysis of the time around an interruption. From this we can conclude that the first five minutes after an interruption and the moment before the next one are the are the phases where productivity is more affected, while the phases around the peak are where the productivity is the highest.
\end{changedforreviewerlong}

@@

In RQ3, we saw different patterns of activity around positive and negative interruptions. In particular, we observed that the edits per minute were higher after a positive interruption; on the contrary they were lower after a negative interruption. The effect is even more marked in the \changedforreviewer{first 5 minutes} after the interruption, as can be seen in the Figure \ref{fig:box_recovery_effect_udc}.

@@

\begin{changedforreviewerlong}
A first challenge that we encountered was that, similarly to the earlier research questions, the high number of different commands made it hard to infer broader trends. We hence created an additional classification of the commands we encountered in the activity logs. In a similar manner to the first classification, we followed an iterative process involving two of the authors to classify the events in these categories. Differently to the first classification, we did not limit ourselves to a restricted number of categories: rather, we let the classification emerge from the iterative process. We ended up with 11 categories, as described in Table \ref{tbl:new_events_udc}. We assigned the label and description that better suited the events grouped on that category and the associated type that reflects the majority of those events.
\end{changedforreviewerlong}

@@

\begin{table}[ht!]
	\small
	\renewcommand{\arraystretch}{1.3}
	\caption{New category of events. }
	\label{tbl:new_events_udc}
	\centering
	\begin{tabular}{|p{2.5cm}|p{9 cm}| p{2cm} |}
		\hline 
		category & description & \changedforreviewer{type} \\  
		\hline 
		\hline 
		high-nav & High-level navigation through classes, windows and artifacts & navigation \\
		file & Management of files (saving, closing, opening, etc.) & navigation\\
		edit-text & Text edition events & edition \\
		debug & Events executed during debugging sessions & navigation \\
		text-nav & Low-level text navigation & edition \\
		search & Text or code search & navigation \\
		find-replace & Search and replacement of text or code entities & navigation\\
		perspective & Change of perspectives in Eclipse & navigation  \\
		control & Source control events & navigation \\
		refactoring & Refactoring events & edition \\
		clean-build &  Events related to the build system (clean, build or run a project)  & navigation \\
		\hline
	\end{tabular}
\end{table}

@@

The second group of interruptions seems to be more disruptive as they require more time to reach the average productivity, and the activities are a reflection of the tasks performed during the recovery time as mentioned by Parnin \cite{PR11}, and they can be matched to the categories high-nav, debug and search, which are the common categories after a negative interruption.
\end{changedforreviewerlong}

@@

\newboolean{showchangedforreviewer}

\setboolean{showchangedforreviewer}{true} % toggle to show or hide sentences and paragraphs requested by the reviewers.
